Responsible Machine Learning\\\large{with Interpretable Models, Post-hoc Explanation, and Disparate Impact Testing}

### 2019-11-07 article-information-2019.tex
several updates

ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{\href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{See: ``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate social biases \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.  ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}. The authors make no claim that the interdependent issues of opaqueness, unwanted sociological bias, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g. the fairwashing of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{shokri2019privacy}. However, this text does present an attempt to address the technological aspects of these vexing problems with interpretable models, post-hoc explanation, and DI testing implemented in widely available, free, and open source Python tools. Section \ref{sec:m_and_m} describes methods and materials herein, including simulated and collected training datasets, interpretable and constrained architectures used to train highly transparent models, post-hoc explanations used to create an \textit{appealable} decision-making framework, tests for DI and other unwanted sociological bias, and public and open source software resources. In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models and post-hoc explanation and DI testing results are also presented. Section \ref{sec:disc} then discusses some nuances of the presented modeling, explanation, and DI testing methods and results. Finally, Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML and also touches on the authors' future work.

@@

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software Resources}\label{ssec:soft}
% Description of code

% github repo, main packges

>>>>

ML models can be inaccurate and unappealable black-boxes, even with the application of newer post-hoc explanation techniques \cite{please_stop}.\footnote{\href{https://www.nytimes.com/2017/06/13/opinion/how-computers-are-harming-criminal-justice.html}{See: ``When a Computer Program Keeps You in Jail''}.} ML models can perpetuate and exacerbate social biases \cite{fairmlbook}, \cite{feldman2015certifying}, \cite{dwork2012fairness}, \cite{gender_shades}.  ML models can be hacked, resulting in manipulated model outcomes or the exposure of proprietary intellectual property or sensitive training data \cite{security_of_ml}, \cite{model_stealing}, \cite{membership_inference}. The authors make no claim that the interdependent issues of opaqueness, unwanted sociological bias, or security vulnerabilities in ML have been solved, even as singular entities, much less as complex intersectional phenomena, e.g. the fairwashing or scaffolding of biased models with ML explanations or the privacy harms of ML explanations \cite{fair_washing}, \cite{scaffolding}, \cite{shokri2019privacy}. However, this text does present an attempt to address the technological aspects of these vexing problems with interpretable models, post-hoc explanation, and DI testing implemented in widely available, free, and open source Python tools. Section \ref{sec:m_and_m} describes methods and materials herein, including simulated and collected training datasets, interpretable and constrained architectures used to train highly transparent models, post-hoc explanations used to create an \textit{appealable} decision-making framework, tests for DI and other unwanted sociological bias, and public and open source software resources. In Section \ref{sec:res}, interpretable and constrained modeling results are compared to less interpretable and unconstrained models and post-hoc explanation and DI testing results are also presented. Section \ref{sec:disc} then discusses some nuances of the presented modeling, explanation, and DI testing methods and results. Finally, Section \ref{sec:con} closes this text with a brief outline of proposed additional steps to increase human trust and understanding in ML and also touches on the authors' future work.

@@

Results are presented for simulated data with known feature importance and social bias characteristics and for more realistic lending data sourced from the Home Mortgage Disclosure Act (HMDA) database.\footnote{\href{https://www.consumerfinance.gov/data-research/hmda/}{See: Mortgage data (HMDA)}.}

\subsubsection{Simulated Data}

\subsubsection{Lending Data}

Lending data is sampled from the public HMDA database to create a real-world dataset on which to assess interpretable models, post-hoc explanations, and DI tests. The sampled lending data contains recent and anonymized consumer loans issued by many major United States (US) financial institutions. Consumers and loans in the dataset are characterized by Bayesian improved surname geocoding (BISG) demographic features, consumer credit scores and various consumer financial features, and loan purpose, amount, and term features.\footnote{\href{https://files.consumerfinance.gov/f/201409_cfpb_report_proxy-methodology.pdf}{See: Using publicly available information to proxy for unidentified race and ethnicity}.} The lending data is divided randomly into training and test partitions. The training data contains 33 total features and 144,000 rows, each representing a unique loan, and a fold identifier to ensure consistent 5-fold cross-validation accuracy and error measurements across different types of models. Consumer financials and loan descriptors are used for training. Demographic features are not used in model training. The lending test data contains 36,000 loans.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software Resources}\label{ssec:soft}
% Description of code
% github repo, main packges
Python code to reproduce the results presented in this text are available at: \url{https://github.com/h2oai/article-information-2019}. The authors primarily make use of the \href{https://github.com/h2oai/datatable}{datatable}, \href{https://github.com/h2oai/h2o-3}{h2o}, \href{https://matplotlib.org/}{matplotlib}, \href{https://pandas.pydata.org/}{pandas}, \href{https://scikit-learn.org/stable/}{scikit-learn}, \href{https://seaborn.pydata.org/}{seaborn}, and \href{https://github.com/slundberg/shap}{shap} packages for data manipulation, modeling, and reporting results.
